Cybersecurity Is Too Important\\To Be Left To The Specialists

### 2019-03-20 FIE2019.tex
Added comments on what to fix
small tidy

% \author{\IEEEauthorblockN{Tom Crick}
% \IEEEauthorblockA{\textit{School of Education} \\
% \textit{Swansea University}\\
% Swansea, UK \\
% thomas.crick@swansea.ac.uk}
% \and
% \IEEEauthorblockN{James H. Davenport}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University of Bath}\\
% Bath, UK \\
% j.h.davenport@bath.ac.uk}
% \and
% \IEEEauthorblockN{Alastair Irons}
% \IEEEauthorblockA{\textit{Faculty of Technology} \\
% \textit{name of organization (of Aff.)}\\
% Sunderland University, Sunderland, UK\\
% alastair.irons@sunderland.ac.uk}
% \and
% \IEEEauthorblockN{Tom Prickett}
% \IEEEauthorblockA{\textit{School of Computing} \\
% \textit{name of organization (of Aff.)}\\
% Northumbria University, Newcastle upon Tyne, UK \\
% tom.prickett@northumbria.ac.uk}
% }

% compact IEEE author format for more than three authors...
\author{\IEEEauthorblockN{Tom Crick\IEEEauthorrefmark{1}, James
    H. Davenport\IEEEauthorrefmark{2}, Alastair
    Irons\IEEEauthorrefmark{3} and Tom Prickett\IEEEauthorrefmark{4}} 
\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Education, 
Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, 
University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
\IEEEauthorblockA{\IEEEauthorrefmark{3}Faculty of Technology, 
Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
\IEEEauthorblockA{\IEEEauthorrefmark{4}School of Computing, 
Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}

@@

This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one. On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of Cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''. It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in Cybersecurity. On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
% JHD what are these?  Also the next paragraph needs thinking about

@@

It is at least plausible to attribute this difference to differences in the accreditation regimes.
\begin{description}
\item[UK]BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accedited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 Juy 2019) \cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}
\section{Policy Context}

@@

Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a})

@@

\subsection{Android}\label{sec:Android}
\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. Their key finding was this.
\begin{quote}
We found that 15.4\% of all 1.3 million Android applications
contained security-related code snippets from
Stack Overflow. Out of these 97.9\% contain at least one
insecure code snippet.
\end{quote}
We should note two caveats (in opposite directions). Their labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure. On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
\par
The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}. Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''. 
\subsection{Agile}
To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for ``Agile'' textbooks. It is the authors' experience that the vast majority of students' individual software projects are developed on Agile lines. Many authorshave found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
\par
Agile's preference for functionality over non-functional requirements is clearly displayed in practice. \cite{Naiakshinaetal2017a} asked 20 student developers to imagine they
were part of a team working on creating a social networking site for
our university and to implement a password storage mechanism for this. 10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not. None of the unprimed ones implemented any security.
\subsection{Staff}
It is well known that Cybersecurity skills are in short supply, for example \cite{Page2018a}.
\begin{quote}
% JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
Research into the state of IT conducted annually by ESG\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years. In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills. This had climbed to 51\% at the beginning of this year. Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
\end{quote}
The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
%JHD possibly to write about his recruitment experiences.  ? more IoC input?

@@

The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK. This could potentially include Cybersecurity related skills. Like the Tech Partnership the focus is upon job market requirements.  Given the size of this initiative this potentially has a significant role to play however at time of writing it is not clear what that shall be.

@@

In recent years the BCS has evolved its Accreditation Practices to promote and mandate the inclusion of Cybersecurity within the programmes the body accredits. The timeline which this process has followed in Table \ref{table:1}.
\begin{center}
  \begin{table}[h!]
  \begin{tabular}{ | p{6cm} |p{1.5cm} |}
    \hline
    IV. ACCREDITATION:RQ3 - A. Developing Expectations &   \\ \hline
    UK Government Cybersecurity Strategy \cite{UKCabinetOffice} & November 2011 \\ \hline
    Three workshops of a consortium of industry, academia and government bodies - led by CPHC and (ISC)2 - leading to the development of Cybersecurity learning guidelines to be embedded into BCS accredited UK Computer Science and IT-related degree \cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
    UK Government report Cybersecurity Skills, Business Perspectives and Government's Next Steps Report Released \cite{UKCabinetOffice2014} & March 2014  \\ \hline
    Council of Professors and Heads of Computing (CPHC) Identifies Cybersecurity as one the top 3 concerns in Computing & April 2014 \\ \hline
    Joint Development of White Paper from CPHC and The International Information Systems Security Certification Consortium(ISC)2 \cite{CPHCISC2014} & April -November 2014 \\ \hline
    Extended Cybersecurity Criteria included in BCS Accreditation Guidelines \cite{BCS2018a}& June 2015 \\
    \hline
    IV. ACCREDITATION:RQ3 - B. What does the BCS tell Universities? & \\ \hline
    Cybersecurity Principles Roadshow & March-April 2016 \\ \hline
    IV. ACCREDITATION:RQ3 - C. Accreditation - what progress has been made? &  \\ \hline
    All visited HEIs expected to be fully compliant & September 2020 (BCS follow a 5-year Accreditation Cycle)  \\ \hline
  \end{tabular}
  \caption{Timeline of the development of Cybersecurity expectations}
  \label{table:1}
  \end{table}
\end{center}
\subsection{Developing Expectations}
Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis. Like many Governments, the UK Goverment has been actively been seeking ways to address this \cite{UKCabinetOffice}, \cite{UKCabinetOffice2014}. In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}. In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of Computer Science Departments) and the Cabinet Office to examine embedding Cybersecurity into undergraduate degrees in the UK. Three workshops in 2013 and 2015 attempted to define the principles of Cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula. Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate Computing Science degrees. This work intially lead to a white paper related to a proposal in the form of a whitepaper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}. The BCS agreed to adopt the outputs into their accreditation criteria. This was the first time that Cybersecurity has been extensively referenced within accreditation criteria for computing and IT-related degrees. The fact that Cybersecurity is included as a  component of the BCS accreditation criteria, reflects the importance placed on Cybersecurity and the expectation that all computing graduates should have knowledge and skills in Cybersecurity as they move towards Chartered status.

@@

\section{Conclusions}
As regards our research questions, we can make the following comments.
\begin{description}
\item[RQ1]The guidelines from both ACM and BCS are good for general education. However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}--) resources. 
\item[RQ2]The recommendation in \cite[p. 98]{ACM2013a} that Cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea. Indeed, a recnt discussion on Cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.  However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.  It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than tthat they not be taught at all.
\par
We commend the BCS-identified good practice of exploring Cybersecurity via a project, possibly in PCI DSS (\S\ref{sec:PCIDSS}).
\item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.
\end{description}

>>>>

% \author{\IEEEauthorblockN{Tom Crick}
% \IEEEauthorblockA{\textit{School of Education} \\
% \textit{Swansea University}\\
% Swansea, UK \\
% thomas.crick@swansea.ac.uk}
% \and
% \IEEEauthorblockN{James H. Davenport}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University of Bath}\\
% Bath, UK \\
% j.h.davenport@bath.ac.uk}
% \and
% \IEEEauthorblockN{Alastair Irons}
% \IEEEauthorblockA{\textit{Faculty of Technology} \\
% \textit{name of organization (of Aff.)}\\
% Sunderland University, Sunderland, UK\\
% alastair.irons@sunderland.ac.uk}
% \and
% \IEEEauthorblockN{Tom Prickett}
% \IEEEauthorblockA{\textit{Department of Computing and Information Sciences} \\
% \textit{name of organization (of Aff.)}\\
% Northumbria University, Newcastle upon Tyne, UK \\
% tom.prickett@northumbria.ac.uk}
% }

% compact IEEE author format for more than three authors...
\author{\IEEEauthorblockN{Tom Crick\IEEEauthorrefmark{1}, James
    H. Davenport\IEEEauthorrefmark{2}, Alastair
    Irons\IEEEauthorrefmark{3} and Tom Prickett\IEEEauthorrefmark{4}} 
\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Education, 
Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, 
University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
\IEEEauthorblockA{\IEEEauthorrefmark{3}Faculty of Technology, 
Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Computer and Information Sciences, 
Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}

@@

This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one. On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of Cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''. It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in Cybersecurity. On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
% JHD what are these?  Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.

@@

It is at least plausible to attribute this difference to differences in the accreditation regimes.
\begin{description}
\item[UK]BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accedited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019) \cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}
\section{Policy Context}

@@

Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.

@@

\subsection{Android}\label{sec:Android}
\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. Their key finding was this.
\begin{quote}
We found that 15.4\% of all 1.3 million Android applications
contained security-related code snippets from
Stack Overflow. Out of these 97.9\% contain at least one
insecure code snippet.
\end{quote}
We should note two caveats (in opposite directions). Their labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure. On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
\par
The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}. Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''. 
\subsection{Agile}
To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for ``Agile'' textbooks. It is the authors' experience that the vast majority of students' individual software projects are developed on Agile lines. Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
\par
Agile's preference for functionality over non-functional requirements is clearly displayed in practice. \cite{Naiakshinaetal2017a} asked 20 student developers to imagine they
were part of a team working on creating a social networking site for
our university and to implement a password storage mechanism for this. 10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not. None of the unprimed ones implemented any security.
\subsection{Staff}
It is well known that Cybersecurity skills are in short supply, for example \cite{Page2018a}.
\begin{quote}
% JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
Research into the state of IT conducted annually by ESG\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years. In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills. This had climbed to 51\% at the beginning of this year. Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
\end{quote}
The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
%JHD possibly to write about his recruitment experiences.  ? more IoC input?

@@

The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK \cite{Davenportetal2019a}. This is likely to include Cybersecurity related skills. Like the Tech Partnership the focus is upon job market requirements. Additionally the Institute is looking at potentially endorsing the demonstrable capabilities of graduates as shown by their university studies, work experience and work placements. Given the size of this initiative this potentially has a significant role to play however at time of writing it is clearly a work in progress.

@@

In recent years the BCS has evolved its Accreditation Practices to promote and mandate the inclusion of Cybersecurity within the programmes the body accredits. The timeline which this process has followed in Table \ref{table:1}.
\begin{center}
  \begin{table}[h!]
  \begin{tabular}{ | p{6cm} |p{1.5cm} |}
    \hline
    IV. ACCREDITATION:RQ3 - A. Developing Expectations &   \\ \hline
    UK Government Cybersecurity Strategy \cite{UKCabinetOffice} & November 2011 \\ \hline
    Three workshops of a consortium of industry, academia and government bodies - led by CPHC and (ISC)2 - leading to the development of Cybersecurity learning guidelines to be embedded into BCS accredited UK Computer Science and IT-related degree \cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
    UK Government report Cybersecurity Skills, Business Perspectives and Government's Next Steps Report Released \cite{UKCabinetOffice2014} & March 2014  \\ \hline
    Council of Professors and Heads of Computing (CPHC) Identifies Cybersecurity as one the top 3 concerns in Computing & April 2014 \\ \hline
    Joint Development of White Paper from CPHC and The International Information Systems Security Certification Consortium(ISC)2 \cite{CPHCISC2014} & April -November 2014 \\ \hline
    Extended Cybersecurity Criteria included in BCS Accreditation Guidelines \cite{BCS2018a}& June 2015 \\
    \hline
    IV. ACCREDITATION:RQ3 - B. What does the BCS tell Universities? & \\ \hline
    Cybersecurity Principles Roadshow & March-April 2016 \\ \hline
    IV. ACCREDITATION:RQ3 - C. Accreditation - what progress has been made? &  \\ \hline
    All visited HEIs expected to be fully compliant & September 2020 (BCS follow a 5-year Accreditation Cycle)  \\ \hline
  \end{tabular}
  \caption{Timeline of the development of Cybersecurity expectations}
  \label{table:1}
  \end{table}
\end{center}
\subsection{Developing Expectations}
Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis. Like many Governments, the UK Goverment has been actively been seeking ways to address this \cite{UKCabinetOffice}, \cite{UKCabinetOffice2014}. In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}. In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of Computer Science Departments) and the Cabinet Office to examine embedding Cybersecurity into undergraduate degrees in the UK. Three workshops in 2013 and 2015 attempted to define the principles of Cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula. Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate Computing Science degrees. This work initially lead to a white paper related to a proposal in the form of a whitepaper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}. The BCS agreed to adopt the outputs into their accreditation criteria. This was the first time that Cybersecurity has been extensively referenced within accreditation criteria for computing and IT-related degrees. The fact that Cybersecurity is included as a  component of the BCS accreditation criteria, reflects the importance placed on Cybersecurity and the expectation that all computing graduates should have knowledge and skills in Cybersecurity as they move towards Chartered status.

@@

\section{Conclusions}
As regards our research questions, we can make the following comments.
\begin{description}
\item[RQ1]The guidelines from both ACM and BCS are good for general education. However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}--) resources. 
\item[RQ2]The recommendation in \cite[p. 98]{ACM2013a} that Cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea. Indeed, a recent discussion on Cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.  However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.  It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than that they not be taught at all.
\par
We commend the BCS-identified good practice of exploring Cybersecurity via a project, possibly in PCI DSS (\S\ref{sec:PCIDSS}).
\item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.
\end{description}
